Flaring star-planet interactions in Kepler and TESS

### 2023-05-23 src/tex/ms.tex
[showyourwork] overleaf sync

% Author list
\author{Ekaterina Ilin}
\author{Katja Poppenh\"ager}
\author{Nikoleta Ilic}

% Abstract with filler text
\begin{abstract}
    In many of the star-planet systems discovered so far, the innermost planets orbit extremely close to their host, that is, within its extended magnetosphere. In these systems, the planets become in-situ probes of the stellar magnetic fields. As they move, they may trigger flares on the stellar surface. Until recently, searches for such flaring star-planet interactions suffered from small sample sizes, and poor orbital phase coverage.

    In this work, we used the excellent phase coverage of Kepler and TESS to search their entire archives for star-planet systems with flaring star-planet interactions. We compiled a catalog of all star-planet systems with flares, and searched for planet-induced flares, i.e., flares clustering in phase with the planet's orbit.

    In the 1811 searched systems, we found 26 single stars with three or more flares each. We quantified the significance of their excess flaring, and compared it against the theoretically expected amplitude of magnetic interaction. Most systems are neither expected to, nor do they show excess flaring. Among the systems that we can expect to interact, we see two branches. An inactive one, where no signs of excess flaring appear, and an active one, where expected power and excess flaring are correlated. This trend is tentative, but consistent for all scaling laws for magnetic interaction.

    Finally, we searched the same data for signs of locally expressed tidal interaction, that is, excess flaring in phase with the tidal bulges ($P_{\rm orb}/2$). We found a similar trend with different scaling laws of tidal interaction, albeit at even lower significance. Our system-by-system comparison suggests that young systems may interact both tidally and magnetically, whereas old, magnetically inactive systems might mostly interact tidally.
    %magnetic field strength (using $Ro$-$B$ relations), orbital parameters, and planet radius.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. The past two decades revealed that planets can occur in orbits so close to their host star that they could almost be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions (SPI) that emphasizes that they are bidirectional, affecting both planet, and star. Specifically, the effects on the star pose challenges, but also opportunities for our understanding of star-planet systems:

On the one hand, a close-in planet that changes the star's rotation and activity makes it more difficult for us to model the system's evolution, as it affects fundamental stellar parameters. Ages inferred from rotation will be biased, if the star is spun up or spun down relative to stars without close-in companions~\citep{tejadaarevalo2021further, brown2014discrepancies, maxted2015comparison}. Analogously, altered activity levels may mimic an older or younger host than it actually is~\citep{ilic2022tidal}. On the other hand, measuring in what ways and how much close-in planets alter their hosts' behavior can provide important insights into the habitability of the entire system~\citep{shkolnik2018signatures}. Even if they orbit far inside the habitable zone, such planets probe the system's space weather and mass loss by revealing the extent of the Alfv\'en zone~\citep{kavanagh2021planetinduced, chebly2022destination}. % Prospectively, planet-induced stellar activity could quantify otherwise difficult to measure planetary properties such as the strength and extent of their magnetic field~\citep{cauley2019magnetic}.

@@

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic star-planet interactions can only occur if the planet is orbiting within the Alfv\'en radius of the star. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby}, and \citet{zarka2007plasma} and \citet{saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. They all lack observational constraints for calibration, which we aim to provide in this work.

@@

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been catalogued~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Currently, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. Their key advantage for this work is the nearly uninterrupted monitoring of star-planet systems that covers all phases of a close-in planet's orbit.

\subsection{Flaring star-planet interactions}
\label{sec:intro:fspi}
The absolute power of flaring star-planet interactions is difficult to constrain. We do not know how energetic planet-induced flares can be. The estimates of the maximum dissipated power~\citep{lanza2018closeby} are well within the regime of superflares typically observed by Kepler and TESS. However, on the one hand, that power is an upper limit, so the observed energies might be lower. The planet might, for instance, be triggering flares prematurely, leading to an overabundance of low energy flares~\citep{loyd2023flares}. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism proposed by~\citep{lanza2012starplanet}, so an individual flare's energy might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should follow the same process as intrinsic flares.

Since planet-induced flares are not expected to look any different from intrinsic flares, and we do not know how their energies will be distributed, we require a method to distinguish them without referring to their individual properties. Instead, we can look at the statistical distribution of flares in time:

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's viewing geometry, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable viewing geometries and magnetic field states, the footpoint visibility will be modulated in phase with the planetary orbit. The resulting phase preference of planet-induced flares introduces a deviation from a random distribution of flares. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

\subsection{Searching for local changes in activity}
\label{sec:intro:local}
In contrast with global SPI, local SPI effects are easier to tell apart from intrinsic stellar behavior because they are tied to the planet's orbital period. 
Phase-correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an X-ray active Hot Jupiter host, the observed slight clustering of flaring in orbital phase remains tentative~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed an eruption of X-ray emission when the Jupiter-sized companion was in periastron on its highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner TRAPPIST-1 planets to look for clustered flaring in phase, but found none. \citet{ilin2022searching} and \citet{klein2022one} found hints of excess flaring, and chromospheric Helium emission in the young AU Mic system, respectively. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

\subsection{Overview}
In this work, we searched for excess flaring in all available Kepler and TESS short cadence light curves of confirmed star-planet systems~(Section~\ref{sec:data}). We scoured the light curves for flares, and tested their observed phases for departures from a uniform distribution in phase with the innermost planet's orbit~(Section~\ref{sec:methods}). We present the catalog of all flares found in our sample of star-planet systems, and compare the test results with the expected power of interaction for each system~(Section~\ref{sec:results}). We discuss the results for the most interesting star-planet systems individually~(Section~\ref{sec:results:individualstars}), and take a look at the patterns emerging from our analysis~(Section~\ref{sec:discussion}), particularly at the possible role of tidal flaring SPI in our observations~(Section~\ref{sec:discussion:tidal}).  We summarize our findings in Section~\ref{sec:summary}.

@@

\section{Data}
\label{sec:data}
In this study, we compiled the largest possible sample to search for flaring star-planet interactions. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the amount of excess flaring from star-planet interaction, we then sought out the orbital periods of each system from the literature~(Section~\ref{sec:data:orbitalperiod}). To compare our measurements to theory, we estimated the expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we also mostly found in the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date results. The parameters for the systems in the final sample of systems are shown in Table~\ref{tab:maintable_lit}.

@@

\subsection{Star-planet systems}
\label{sec:data:sps}
We compiled our sample of star-planet systems from the PSCP, from which we removed controversial planet detections (``pl\_controv\_flag'' set to 1), i.e., detections with existing literature questioning the result. The resulting catalog contained 2993 transiting and 191 non-transiting systems, from which we picked the innermost known planet in each system.

\subsection{Kepler and TESS photometry}
\label{sec:data:photometry}
The Kepler and TESS missions are unbeaten with respect to long-term optical monitoring of stellar flares. Their excellent coverage of orbital phases makes the light curves ideal for our search for orbital phase dependent flare timing. Between 2009 and 2013, the Kepler space telescope~\citep{koch2010kepler} nearly continuously observed a patch of the sky in the Cygnus-Lyra region. Each of the 18 observing Quarters contains nearly uninterrupted $\sim 90$ days of observations, totaling over $100\,000$ stars monitored in 2-min cadence in a broad $400-850\,$nm bandpass.
% 100000 from here: https://exoplanetarchive.ipac.caltech.edu/docs/KeplerMission.html

The Transiting Exoplanet Survey Satellite~(TESS,~\cite{ricker2015transiting}) is an all-sky mission that began operations in 2018, completed two sky scans until summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the $600-1000\,$nm band for $\sim 27\,$d in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations alone, with about $20,000$ targets per Sector. Out of these, from Sector 27 on, $1,000$ targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

Based on the filtered PSCP table, we use the \texttt{lightkurve}~\citep{lightkurvecollaboration2018lightkurve} software to query the full Kepler archive (quarters Q0-Q17, Data Release 25), and the most recent TESS catalog (July 2022) for their respective 1-min and 2-min/20-s cadence light curves. In total, we searched 7258 light curves for flares -- 3030 Kepler Quarters, and 4228 TESS Sectors. We found and analyzed light curves for a total of 1811 systems, 344 of which were observed only by the primary Kepler mission, 1205 only by TESS, and 262 by both missions. We did not use K2 light curves because of the many systematics, and consequently the high time investment of de-trending them. But we included the flare list obtained by~\citet{paudel2018k2} for TRAPPIST-1 K2 flares, since this system is a potential candidate for sub-Alfv\'enic interactions~\citep{fischer2019timevariable}.

\subsection{Orbital periods and transit mid-times}
\label{sec:data:orbitalperiod}
We adopt the orbital periods from the PSCP, which were either obtained from Kepler or TESS transits, or from radial velocity measurements. For the transiting systems, we use the transit mid-times given in the PSCP to set orbital phase zero. If possible, we use the transit mid-times determined in each mission separately to reduce the uncertainties on orbital phase. For the non-transiting system, phase zero is set arbitrarily.
\subsection{Semi-major axes and eccentricities}
\label{sec:data:a}
The critical parameter is the distance between the star and the planet, not the semi-major axis itself. This varies if the orbit is eccentric. To account for varying distance, we adopt the mean semi-major axes from the literature, but use a custom estimate for the uncertainty that includes the eccentricity.

@@

Among the stars without given uncertainties, the rotation of GJ 3082 was measured and is consistent both in TESS and in KELT light curves to below 1\,s precision~\citep{gunther2020stellar}. The rotation of GJ 674 is consistent between light curve periodograms~\citep{kiraga2007agerotationactivity}, and activity-rotation relations~\citep{boisse2011disentangling} at a $<10\%$ level. GJ 3323 only has rotation periods inferred from activity indicators~\citep{Astudillo-defru2017magnetic}, without any periodicity detected in the activity indicators themselves~\citep{astudillo-defru2017harps}. The rotation of Proxima Cen is very slow, but consistent between different datasets~\citep{anglada-escude2016terrestrial, kiraga2007agerotationactivity}. In general, among the stars without given uncertainties on the rotation period, the period is either 
\begin{itemize}
    \item short, leaving a clear peak in the periodogram, and little doubt about the period after ruling out aliases with $P/2$, as with GJ 3082, or
    \item longer than the Sun's, that is, in the regime where rotation-activity relations are reliable because the stars are in the unsaturated activity regime, as in the case of GJ 674.
\end{itemize}

@@

The main data for this analysis are flare times. To obtain them, we gather the Kepler and TESS light curves for all star-planet systems, remove rotational variability trends, search the de-trended light curves for flares, and convert the flare times to orbital phases. We then perform a customized Anderson-Darling test on the flare peak time distribution, which yields a p-value for the significance of the SPI signal. Finally, we compare these results to the theoretically expected SPI power $P$ in each system.

The methods for light curve de-trending and flare finding in Kepler and TESS light curves, as well as the Anderson-Darling test, are the same as detailed in~\citet{ilin2022searching}. We briefly recap the techniques in Sections~\ref{sec:methods:flaresearch} and \ref{sec:methods:adtest}. The expected power of SPI depends on the relative velocity between the planet and the magnetic field strength in its orbit, the derivation of which we explain in 
Sections~\ref{sec:methods:relvel} and \ref{sec:methods:bfield}, respectively. We can then combine them with stellar radius, planetary radius, and semi-major axis to estimate the power of SPI. We use the scaling laws from two different magnetic SPI models, which we introduce in Section~\ref{sec:methods:pspi}.

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process, implemented as the \texttt{custom\_detrending} method in \texttt{AltaiPony}, a flare science package for light curve analysis~\citep{ilin2021altaipony}. First, we apply a spline fit with a coarse sampling of 30h-averaged values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\citep{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ (or $1.5 \sigma$ for very active stars like AU Mic or Proxima Cen) threshold as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked. Transits are usually too shallow to affect either light curve de-trending or flare finding. The light curve portions around deep transits were inspected manually.

In the de-trended light curves, we search for flares as series of at least three consecutive data points $3\sigma$ above the noise level using the \texttt{AltaiPony} method \texttt{find\_flares}. We estimate the noise level as the standard deviation in a rolling window of two hours, while masking deviation above $2.5\sigma$ (or $1.5$ sigma for active stars). To capture the exponential tail of the flare, we use the \texttt{addtail} flag in the \texttt{find\_flares} method to continue adding data points to the end of the flare as long as they are $2\sigma$ above the noise threshold.

We vet all flare candidates by eye, and exclude instrumental false positives and physical one like Solar System Objects. For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the integrated flare flux $F_{flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{flare}(t)}{F_0}.
\end{equation}
Equivalent duration is the time during which the non-flaring star releases as much energy as the flare.

>>>>

% Author list
\author{Ekaterina Ilin}
\author{Katja Poppenh\"ager}
\author{Judy Chebly}
\author{Nikoleta Ilic}

% Abstract with filler text
\begin{abstract}
    In many of the star-planet systems discovered so far, the innermost planets orbit within only a few stellar radii. In these systems, the planets become in-situ probes of the extended stellar magnetic field. As they move, they can trigger flares on the stellar surface. Until recently, however, searches for such planet-induced flaring suffered from small sample sizes, and poor orbital phase coverage.

    In this work, we used the excellent phase coverage of the archives of Kepler and the Transiting Exoplanet Survey Satellite to search for star-planet systems with flaring star-planet interactions. We compiled a catalog of all their flares, and searched for planet-induced flaring, i.e., flares clustering in phase with the planet's orbit.

    In the 1811 searched systems, we found 25 single stars with three or more flares each. We quantified the significance of their excess flaring, and compared it against the theoretically expected amplitude of magnetic interaction. As we did not select for short orbits, most systems do not show excess flaring, as expected. Those we expect to interact fall on two branches. An inactive one, without any signs of excess flaring, and an active one, where expected power and excess flaring are correlated. This trend is tentative, but consistent for several scaling laws for magnetic interaction.

    Finally, we searched our data for locally expressed tidal interaction, i.e., excess flaring in phase with the tidal bulges ($P_{\rm orb}/2$). We find a marginal trend with different scaling laws of tidal interaction. Our system-by-system comparison tentatively suggests that young systems can interact both tidally and magnetically, whereas old, magnetically inactive systems mostly interact tidally.
    %magnetic field strength (using $Ro$-$B$ relations), orbital parameters, and planet radius.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. The past two decades revealed that planets can occur in orbits of only a few stellar radii~[e.g., ][]\citep{sanchis-ojeda2014study} that they could almost be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions (SPI) that emphasizes that they are bidirectional, affecting both planet, and star. Specifically, the effects on the star pose challenges, but also opportunities for our understanding of star-planet systems:

On the one hand, a close-in planet that changes the star's rotation and activity makes it more difficult for us to model the system's evolution, as it affects fundamental stellar parameters. Ages inferred from rotation will be biased, if the star is spun up or spun down relative to stars without close-in companions~\citep{tejadaarevalo2021further, brown2014discrepancies, maxted2015comparison}. Analogously, altered activity levels may mimic an older or younger host than it actually is~\citep{ilic2022tidal}. On the other hand, measuring in what ways and how much close-in planets alter their hosts' behavior can provide important insights into the habitability of the entire system~\citep{shkolnik2018signatures}. Even if they orbit far inside the habitable zone, such planets probe the system's space weather and mass loss by revealing the extent of the Alfv\'en zone through SPI~\citep{kavanagh2021planetinduced, chebly2022destination}. % Prospectively, planet-induced stellar activity could quantify otherwise difficult to measure planetary properties such as the strength and extent of their magnetic field~\citep{cauley2019magnetic}.

@@

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic star-planet interactions can occur if the planet is orbiting within the Alfv\'en radius of the star at least part of the time~\citep{preusse2006magnetic, cohen2011dynamics}. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby}, and \citet{zarka2007plasma} and \citet{saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. They all lack observational constraints from large samples for calibration, which we aim to provide in this work.

@@

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been catalogued~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Since 2018, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. The key advantage of both missions for this work is the nearly uninterrupted monitoring of star-planet systems that covers all phases of a close-in planet's orbit.

\subsection{Flaring star-planet interactions}
\label{sec:intro:fspi}
The absolute power of flaring star-planet interactions is difficult to constrain. We do not know how energetic planet-induced flares can be. The estimates for dissipated energy~\citep{lanza2018closeby} are in the ballpark of the regime of superflares observed by Kepler and TESS. However, on the one hand, the planet might be triggering flares prematurely, which would be of lower energy~\citep{loyd2023flares}, and end up below our detection threshold. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism proposed by~\citep{lanza2012starplanet}, so an individual flare's energy might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We will assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should still have the same physics as intrinsic flares.

Since planet-induced flares might not look any different from intrinsic flares, and we do not know how their energies will be distributed, we require a method to distinguish them without referring to their individual properties. Instead, we can look at the statistical distribution of flares in time:

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's viewing geometry, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable viewing geometries and magnetic field strengths and configurations, the footpoint visibility will be modulated in phase with the planetary orbit. The resulting phase preference of planet-induced flares introduces a deviation from a random distribution of flares. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

\subsection{Searching for local changes in activity}
\label{sec:intro:local}
In contrast to global SPI, local SPI effects are easier to tell apart from intrinsic stellar behavior because they are tied to the planet's orbital period. 
Phase-correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an active Hot Jupiter host, higher X-ray energy flares seem only marginally clustered in orbital phase than lower energy ones~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed an eruption of X-ray emission when the Jupiter-sized companion was in periastron on the highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner \mbox{TRAPPIST-1} planets to look for clustered flaring in phase, but found none. \citet{ilin2022searching} and \citet{klein2022one} found hints of excess flaring, and chromospheric Helium emission in the young AU Mic system, respectively. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

\subsection{Overview}
In this work, we searched for excess flaring in all available Kepler and TESS short cadence light curves of confirmed star-planet systems~(Section~\ref{sec:data}). We scoured the light curves for flares, and tested their observed phases for departures from a uniform distribution in phase with the innermost planet's orbit~(Section~\ref{sec:methods}). We present the catalog of all flares found in our sample of star-planet systems, and compare the test results with the expected power of interaction for each system~(Section~\ref{sec:results}). We discuss the results for the most interesting star-planet systems individually~(Section~\ref{sec:results:individualstars}), and take a look at the patterns emerging from our analysis~(Section~\ref{sec:discussion}), and the possible role of tidal flaring SPI in our observations~(Section~\ref{sec:discussion:tidal}).  We summarize our findings in Section~\ref{sec:summary}.

@@

\section{Data}
\label{sec:data}
In this study, we aimed at compiling the largest possible sample to search for flaring star-planet interactions. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the amount of excess flaring from star-planet interaction, we further required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). To compare our measurements to theory, we estimated the expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we mostly took from the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date results. The parameters for the systems in the final sample of 25 systems are shown in Table~\ref{tab:maintable_lit}.

@@

\subsection{Star-planet systems}
\label{sec:data:sps}
We compiled our sample of star-planet systems from the PSCP, from which we removed controversial planet detections (``pl\_controv\_flag'' must be 0), i.e., detections with existing literature questioning the result. The resulting catalog contained 2993 transiting and 191 non-transiting systems, from which we picked the innermost known planet in each system.

\subsection{Kepler and TESS photometry}
\label{sec:data:photometry}
The Kepler and TESS missions are unbeaten with respect to long-term optical monitoring of stellar flares. Their excellent coverage of orbital phases makes the light curves ideal for our search for orbital phase dependent flaring. Between 2009 and 2013, the Kepler space telescope~\citep{koch2010kepler} nearly continuously observed a patch of the sky in the Cygnus-Lyra region. Each of the 18 observing Quarters contains nearly uninterrupted $\sim 90$ days of observations, totaling over $100\,000$ stars monitored in 2-min cadence in a broad $400-850\,$nm bandpass.
% 100000 from here: https://exoplanetarchive.ipac.caltech.edu/docs/KeplerMission.html

The Transiting Exoplanet Survey Satellite~[TESS, ][]\citep{ricker2015transiting} is an all-sky mission that began operations in 2018, completed two sky scans until summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the $600-1000\,$nm band for $\sim 27\,$d in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations alone, with about $20,000$ targets per Sector. Out of these, from Sector 27 on, $1,000$ targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

Based on the filtered PSCP table, we use the \texttt{lightkurve}~\citep{lightkurvecollaboration2018lightkurve} Python software to query the full Kepler archive (quarters Q0-Q17, Data Release 25), and the most recent TESS catalog (July 2022) for their respective 1-min and 2-min/20-s cadence light curves. In total, we searched 7258 light curves for flares -- 3030 Kepler Quarters, and 4228 TESS Sectors. We found and analyzed light curves for a total of 1811 systems, 344 of which were observed only by the primary Kepler mission, 1205 only by TESS, and 262 by both missions. We did not use K2 light curves because of the many systematics, and consequently the high time investment of de-trending them. But we included the flare list obtained by~\citet{paudel2018k2} for TRAPPIST-1 K2 flares, since this system is a potential candidate for sub-Alfv\'enic interactions~\citep{fischer2019timevariable}.

\subsection{Orbital periods and transit mid-times}
\label{sec:data:orbitalperiod}
We adopt the orbital periods from the PSCP, which were either obtained from Kepler or TESS transits, or from radial velocity measurements. For the transiting systems, we use the transit mid-times given in the PSCP to set orbital phase zero. If possible, we use the transit mid-times determined in each mission separately to reduce the uncertainties on orbital phase. For the non-transiting system, phase zero is set arbitrarily.
\subsection{Semi-major axes and eccentricities}
\label{sec:data:a}
A critical parameter for the possibility of magnetic SPI and its power is the distance between the star and the planet, not the semi-major axis itself. This varies if the orbit is eccentric. We therefore adopt the mean semi-major axes from the literature, but use a custom estimate for the uncertainty that includes the eccentricity:

@@

Among the stars without given uncertainties, the rotation of GJ 3082 was measured and is consistent both in TESS and in KELT light curves to below 1\,s precision~\citep{gunther2020stellar}. The rotation of GJ 674 is consistent between light curve periodograms~\citep{kiraga2007agerotationactivity}, and activity-rotation relations~\citep{boisse2011disentangling} at a $<10\%$ level. GJ 3323 only has rotation periods inferred from activity indicators~\citep{astudillo-defru2017magnetic}, without any periodicity detected in the activity indicators themselves~\citep{astudillo-defru2017harps}. The rotation of Proxima Cen is very slow, but consistent between different datasets~\citep{anglada-escude2016terrestrial, kiraga2007agerotationactivity}. In general, among the stars without given uncertainties on the rotation period, the period is either 
\begin{itemize}
    \item short, leaving a clear peak in the periodogram, and little doubt about the period after ruling out aliases with $P/2$, as with GJ 3082, or
    \item longer than the Sun's, that is, in the regime where rotation-activity relations are reliable because the stars are in the unsaturated activity regime, as in the case of GJ 674, and GJ 3323.
\end{itemize}

@@

The main data for this analysis are flare times. To obtain them, we gather the Kepler and TESS light curves for all star-planet systems, remove rotational variability trends, search the de-trended light curves for flares, and convert the flare times to orbital phases. We then perform a customized Anderson-Darling test on the flare peak time distribution, which yields a $p$-value for the significance of the SPI signal. Finally, we compare these results to the theoretically expected SPI power $P_{xx}$ in each system.

The methods for light curve de-trending and flare finding in Kepler and TESS light curves, as well as the Anderson-Darling test, are the same as detailed in~\citet{ilin2022searching}. We briefly recap the techniques in Sections~\ref{sec:methods:flaresearch} and \ref{sec:methods:adtest}. The expected power of SPI depends on the relative velocity between the planet and the magnetic field strength in its orbit, the derivation of which we explain in 
Sections~\ref{sec:methods:relvel} and \ref{sec:methods:bfield}, respectively. We can then combine them with stellar radius, planetary radius, and semi-major axis to estimate the power of SPI. We use the scaling laws for two different magnetic SPI models, which we introduce in Section~\ref{sec:methods:pspi}.

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process, implemented as the \texttt{custom\_detrending} method in \texttt{AltaiPony}, a flare science Python package for light curve analysis~\citep{ilin2021altaipony}. First, we apply a spline fit with a coarse sampling of 30h-averaged values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\citep{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ (or $1.5 \sigma$ for stars like AU Mic or Proxima Cen, which are very active or have very low noise levels, or both) threshold as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked. Transits are usually too shallow to affect either light curve de-trending or flare finding. The light curve portions around deep transits were inspected manually.

In the de-trended light curves, we search for flares as series of at least three consecutive data points $3\sigma$ above the noise level using the \texttt{AltaiPony} method \texttt{find\_flares}. We estimate the noise level as the standard deviation in a rolling window of two hours, while masking deviation above $2.5\sigma$ (or $1.5\sigma$  for stars like AU Mic and Proxima Cen). To capture the exponential tail of the flare, we use the \texttt{addtail} flag in the \texttt{find\_flares} method to continue adding data points to the end of the flare as long as they are $2\sigma$ above the noise threshold.

We vet all flare candidates by eye, and exclude instrumental false positives and physical ones like passing Solar System Objects. For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the integrated flare flux $F_{\rm flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}
Equivalent duration is equal to the time during which the non-flaring star releases as much energy as the flare.

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}. In brief, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. We repeat this test with different phase offsets to account for potential biases in sensitivity of the test at different phases.
